An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-06 EconGrowth.Rnw
Added a draft conclusion section.

\noindent Equation \ref{eq:CESe} augments a standard CES production function (Equation \ref{eq:CES}) with energy using a ($kl$)($e$) nesting structure, as is typical in the literature. Equation \ref{eq:CES} is a degenerate form of Equation \ref{eq:CESe} where $\delta = 1$. The elasticity of substitution between capital ($k$) and labor ($l$) is given by $\sigma_1 = \frac{1}{1+\rho_1}$, the elasticity of substitution between ($kl$) and ($e$) is given by $\sigma = \frac{1}{1+\rho}$, $\delta \in [0, 1]$ describes the importance of ($kl$) relative to ($e$), and $\gamma$ is a fitting parameter, expected to be close to unity.

@@

****************** Add conclusion here. *********************

>>>>

\noindent Equation \ref{eq:CES} is a CES production function with capital stock ($k$) and labor ($l$) factors of production. Equation \ref{eq:CESe} augments Equation \ref{eq:CES} with energy using a ($kl$)($e$) nesting structure, as is typical in the literature. Equation \ref{eq:CES} is a degenerate form of Equation \ref{eq:CESe} where $\delta = 1$. The elasticity of substitution between capital ($k$) and labor ($l$) is given by $\sigma_1 = \frac{1}{1+\rho_1}$, the elasticity of substitution between ($kl$) and ($e$) is given by $\sigma = \frac{1}{1+\rho}$, $\delta \in [0, 1]$ describes the importance of ($kl$) relative to ($e$), and $\gamma$ is a fitting parameter, expected to be close to unity.

@@

****************** Edit conclusion after we finish the statistical studies. *********************

We found that statistical properties of historical economic data (e.g., covariances) and statistical properties of fitting parameters in production functions (e.g., confidence intervals) are extremely useful for interpreting historical economic data and growth model results. We know of no previous energy-economic studies that presented these statistical properties, and we presume that energy-augmentation of production functions has been done in the literature without regard to whether it is statistically meaningful to do so. That different authors can find good correlation between model predictions and historical data with very different growth model coefficients is a result of the heretofore unseen large confidence intervals on the fitting parameters. As a consequence, economic growth models that assume very different relationships among capital stock, labor, and energy can all be successful in matching historical data. The production function is a canvas upon which investigators can paint their preconceived vision of the importance (or not) of energy for ecnomic growth.

To break free of this problem, a new way of conceptualizing the relationships among capital stock, labor, and energy is required. Our analysis above indicates that energy is unequivocally complementary to the capital stock/labor pair for many types of economies. We agree with K\"{u}mmel that energy is required to form and maintain capital stock, and that energy is required to sustain labor. (And, for that matter, that labor is required to operate capital stock.) I.e., energy is already always \emph{embedded within} capital stock and labor. We posit that the complementarity between ($kl$) and ($e$) shown by the energy-augmented CES production function is due to the embeddedness of energy within the traditional factors of production, capital stock and labor.

As a consequence, production functions that use only energy and production functions that use the capital stock/labor pair are both successful in predicting economic growth. The augmentation of capital stock/labor production functions with energy is both unwarranted and unnecessary.
